---
title: "Predicting UFC Fight Results"
output:
  html_document:
    df_print: paged
date: '2023-12-11'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(ggplot2)
library(readr)
library(dplyr)
library(knitr)
library(ggplot2)
library(lubridate)
library(purrr)
library(randomForest)

data <- read_csv("data (1).csv")
```

```{r}
data %>%
  group_by(weight_class) %>%
  summarise(count = n())
```

```{r}
data <- data %>%
  mutate(fight_gender = ifelse(weight_class %in% c("WomenBantamweight", "WomenFeatherweight", "WomenFlyweight",
                                             "WomenStrawweight"), "women", "men"))

ggplot(data = data, aes(x = fight_gender)) + geom_bar()
ggplot(data = data, aes(x = Winner)) + geom_bar()
```

```{r}
data <- data %>%
  mutate(first_fight = ifelse(is.na(B_avg_KD) | is.na(R_avg_KD), "yes", "no"))
ggplot(data = data, aes(x = first_fight)) + geom_bar()
```


```{r}
data$date <- as.Date(data$date, format = "%m/%d/%y")

# Extract year from the Date column
data$year <- as.numeric(format(data$date, "%Y"))


year_freq <- data %>% group_by(year) %>% summarize(count = n()) 

ggplot(data = year_freq, mapping = aes(x = year, y = count)) + geom_line() + labs(x = "Year", y = "Number of Fights", title = "Trend of Number of Fights in the UFC")

```

```{r}
data2 <- data %>%
  filter(fight_gender == "men" & first_fight == "no" & weight_class != "CatchWeight")
nrow(data2)
```


#randomizing blue and red
```{r}
data2 <- data2 %>%
  mutate(Winner2 = ifelse(Winner == "Red", R_fighter, B_fighter)) %>%
  rename(B_avg_CTRL_time.seconds. = `B_avg_CTRL_time(seconds)`,
         B_avg_opp_CTRL_time.seconds. = `B_avg_opp_CTRL_time(seconds)`,
         B_total_time_fought.seconds. = `B_total_time_fought(seconds)`,
         B_win_by_KO.TKO = `B_win_by_KO/TKO`,
         R_avg_CTRL_time.seconds. = `R_avg_CTRL_time(seconds)`,
         R_avg_opp_CTRL_time.seconds. = `R_avg_opp_CTRL_time(seconds)`,
         R_total_time_fought.seconds. = `R_total_time_fought(seconds)`,
         R_win_by_KO.TKO = `R_win_by_KO/TKO`)
```


```{r}
shuffle_fighters <- function(B_fighter, B_avg_KD, B_avg_opp_KD, B_avg_SIG_STR_pct, B_avg_opp_SIG_STR_pct,
                             B_avg_TD_pct,B_avg_opp_TD_pct, B_avg_SUB_ATT, 
                             B_avg_opp_SUB_ATT, B_avg_REV, B_avg_opp_REV, 
                             B_avg_SIG_STR_att, B_avg_SIG_STR_landed, 
                             B_avg_opp_SIG_STR_att,B_avg_opp_SIG_STR_landed, 
                             B_avg_TOTAL_STR_att, B_avg_TOTAL_STR_landed,
                             B_avg_opp_TOTAL_STR_att, B_avg_opp_TOTAL_STR_landed, 
                             B_avg_TD_att, B_avg_TD_landed, B_avg_opp_TD_att, 
                             B_avg_opp_TD_landed, B_avg_HEAD_att,
                             B_avg_HEAD_landed, B_avg_opp_HEAD_att, 
                             B_avg_opp_HEAD_landed, B_avg_BODY_att,
                             B_avg_BODY_landed, B_avg_opp_BODY_att, 
                             B_avg_opp_BODY_landed, B_avg_LEG_att, 
                             B_avg_LEG_landed, B_avg_opp_LEG_att, 
                             B_avg_opp_LEG_landed, B_avg_DISTANCE_att,
                             B_avg_DISTANCE_landed, B_avg_opp_DISTANCE_att, 
                             B_avg_opp_DISTANCE_landed,B_avg_CLINCH_att, 
                             B_avg_CLINCH_landed, B_avg_opp_CLINCH_att, 
                             B_avg_opp_CLINCH_landed, B_avg_GROUND_att, B_avg_GROUND_landed,
                             B_avg_opp_GROUND_att, B_avg_opp_GROUND_landed, B_avg_CTRL_time.seconds.,
                             B_avg_opp_CTRL_time.seconds., B_total_time_fought.seconds., 
                             B_total_rounds_fought, B_total_title_bouts, B_current_win_streak, 
                             B_current_lose_streak, B_longest_win_streak, B_wins, B_losses, B_draw,
                             B_win_by_Decision_Majority, B_win_by_Decision_Split,
                             B_win_by_Decision_Unanimous, B_win_by_KO.TKO, B_win_by_Submission,
                             B_win_by_TKO_Doctor_Stoppage, B_Stance, B_Height_cms, B_Reach_cms, B_Weight_lbs,
                             B_age, R_fighter, R_avg_KD, R_avg_opp_KD, R_avg_SIG_STR_pct, 
                             R_avg_opp_SIG_STR_pct, R_avg_TD_pct, R_avg_opp_TD_pct, R_avg_SUB_ATT, 
                             R_avg_opp_SUB_ATT, R_avg_REV, R_avg_opp_REV, R_avg_SIG_STR_att, 
                             R_avg_SIG_STR_landed, R_avg_opp_SIG_STR_att,
                             R_avg_opp_SIG_STR_landed, R_avg_TOTAL_STR_att, R_avg_TOTAL_STR_landed,
                             R_avg_opp_TOTAL_STR_att, R_avg_opp_TOTAL_STR_landed, R_avg_TD_att,
                             R_avg_TD_landed, R_avg_opp_TD_att, R_avg_opp_TD_landed, R_avg_HEAD_att,
                             R_avg_HEAD_landed, R_avg_opp_HEAD_att, R_avg_opp_HEAD_landed, R_avg_BODY_att,
                             R_avg_BODY_landed, R_avg_opp_BODY_att, R_avg_opp_BODY_landed, R_avg_LEG_att,
                             R_avg_LEG_landed, R_avg_opp_LEG_att, R_avg_opp_LEG_landed, R_avg_DISTANCE_att,
                             R_avg_DISTANCE_landed, R_avg_opp_DISTANCE_att, R_avg_opp_DISTANCE_landed,
                             R_avg_CLINCH_att, R_avg_CLINCH_landed, R_avg_opp_CLINCH_att,
                             R_avg_opp_CLINCH_landed, R_avg_GROUND_att, R_avg_GROUND_landed,
                             R_avg_opp_GROUND_att, R_avg_opp_GROUND_landed, R_avg_CTRL_time.seconds.,
                             R_avg_opp_CTRL_time.seconds., R_total_time_fought.seconds.,
                             R_total_rounds_fought, R_total_title_bouts, R_current_win_streak,
                             R_current_lose_streak, R_longest_win_streak, R_wins, R_losses, R_draw,
                             R_win_by_Decision_Majority, R_win_by_Decision_Split,
                             R_win_by_Decision_Unanimous, R_win_by_KO.TKO, R_win_by_Submission,
                             R_win_by_TKO_Doctor_Stoppage, R_Stance, R_Height_cms, R_Reach_cms,
                             R_Weight_lbs, R_age) {
  if (runif(1) < 0.5) {
    return(c(B_fighter, B_avg_KD, B_avg_opp_KD, 
                             B_avg_SIG_STR_pct, B_avg_opp_SIG_STR_pct, 
                             B_avg_TD_pct,B_avg_opp_TD_pct, B_avg_SUB_ATT, 
                             B_avg_opp_SUB_ATT, B_avg_REV, B_avg_opp_REV, 
                             B_avg_SIG_STR_att, B_avg_SIG_STR_landed, 
                             B_avg_opp_SIG_STR_att,B_avg_opp_SIG_STR_landed, 
                             B_avg_TOTAL_STR_att, B_avg_TOTAL_STR_landed,
                             B_avg_opp_TOTAL_STR_att, B_avg_opp_TOTAL_STR_landed, 
                             B_avg_TD_att, B_avg_TD_landed, B_avg_opp_TD_att, 
                             B_avg_opp_TD_landed, B_avg_HEAD_att,
                             B_avg_HEAD_landed, B_avg_opp_HEAD_att, 
                             B_avg_opp_HEAD_landed, B_avg_BODY_att,
                             B_avg_BODY_landed, B_avg_opp_BODY_att, 
                             B_avg_opp_BODY_landed, B_avg_LEG_att, 
                             B_avg_LEG_landed, B_avg_opp_LEG_att, 
                             B_avg_opp_LEG_landed, B_avg_DISTANCE_att,
                             B_avg_DISTANCE_landed, B_avg_opp_DISTANCE_att, 
                             B_avg_opp_DISTANCE_landed,B_avg_CLINCH_att, 
                             B_avg_CLINCH_landed, B_avg_opp_CLINCH_att, 
                             B_avg_opp_CLINCH_landed, B_avg_GROUND_att, B_avg_GROUND_landed,
                             B_avg_opp_GROUND_att, B_avg_opp_GROUND_landed, B_avg_CTRL_time.seconds.,
                             B_avg_opp_CTRL_time.seconds., B_total_time_fought.seconds., 
                             B_total_rounds_fought, B_total_title_bouts, B_current_win_streak, 
                             B_current_lose_streak, B_longest_win_streak, B_wins, B_losses, B_draw,
                             B_win_by_Decision_Majority, B_win_by_Decision_Split,
                             B_win_by_Decision_Unanimous, B_win_by_KO.TKO, B_win_by_Submission,
                             B_win_by_TKO_Doctor_Stoppage, B_Stance, B_Height_cms, B_Reach_cms, B_Weight_lbs,
                             B_age, R_fighter, R_avg_KD, R_avg_opp_KD, R_avg_SIG_STR_pct, 
                             R_avg_opp_SIG_STR_pct, R_avg_TD_pct, R_avg_opp_TD_pct, R_avg_SUB_ATT, 
                             R_avg_opp_SUB_ATT, R_avg_REV, R_avg_opp_REV, R_avg_SIG_STR_att, 
                             R_avg_SIG_STR_landed, R_avg_opp_SIG_STR_att,
                             R_avg_opp_SIG_STR_landed, R_avg_TOTAL_STR_att, R_avg_TOTAL_STR_landed,
                             R_avg_opp_TOTAL_STR_att, R_avg_opp_TOTAL_STR_landed, R_avg_TD_att,
                             R_avg_TD_landed, R_avg_opp_TD_att, R_avg_opp_TD_landed, R_avg_HEAD_att,
                             R_avg_HEAD_landed, R_avg_opp_HEAD_att, R_avg_opp_HEAD_landed, R_avg_BODY_att,
                             R_avg_BODY_landed, R_avg_opp_BODY_att, R_avg_opp_BODY_landed, R_avg_LEG_att,
                             R_avg_LEG_landed, R_avg_opp_LEG_att, R_avg_opp_LEG_landed, R_avg_DISTANCE_att,
                             R_avg_DISTANCE_landed, R_avg_opp_DISTANCE_att, R_avg_opp_DISTANCE_landed,
                             R_avg_CLINCH_att, R_avg_CLINCH_landed, R_avg_opp_CLINCH_att,
                             R_avg_opp_CLINCH_landed, R_avg_GROUND_att, R_avg_GROUND_landed,
                             R_avg_opp_GROUND_att, R_avg_opp_GROUND_landed, R_avg_CTRL_time.seconds.,
                             R_avg_opp_CTRL_time.seconds., R_total_time_fought.seconds.,
                             R_total_rounds_fought, R_total_title_bouts, R_current_win_streak,
                             R_current_lose_streak, R_longest_win_streak, R_wins, R_losses, R_draw,
                             R_win_by_Decision_Majority, R_win_by_Decision_Split,
                             R_win_by_Decision_Unanimous, R_win_by_KO.TKO, R_win_by_Submission,
                             R_win_by_TKO_Doctor_Stoppage, R_Stance, R_Height_cms, R_Reach_cms,
                             R_Weight_lbs, R_age))
  } else {
    return(c(R_fighter, R_avg_KD, R_avg_opp_KD, 
                             R_avg_SIG_STR_pct, R_avg_opp_SIG_STR_pct, 
                             R_avg_TD_pct, R_avg_opp_TD_pct, R_avg_SUB_ATT, 
                             R_avg_opp_SUB_ATT, R_avg_REV, R_avg_opp_REV, 
                             R_avg_SIG_STR_att, R_avg_SIG_STR_landed, 
                             R_avg_opp_SIG_STR_att, R_avg_opp_SIG_STR_landed, 
                             R_avg_TOTAL_STR_att, R_avg_TOTAL_STR_landed,
                             R_avg_opp_TOTAL_STR_att, R_avg_opp_TOTAL_STR_landed, 
                             R_avg_TD_att, R_avg_TD_landed, R_avg_opp_TD_att, 
                             R_avg_opp_TD_landed, R_avg_HEAD_att,
                             R_avg_HEAD_landed, R_avg_opp_HEAD_att, 
                             R_avg_opp_HEAD_landed, R_avg_BODY_att,
                             R_avg_BODY_landed, R_avg_opp_BODY_att, 
                             R_avg_opp_BODY_landed, R_avg_LEG_att, 
                             R_avg_LEG_landed, R_avg_opp_LEG_att, 
                             R_avg_opp_LEG_landed, R_avg_DISTANCE_att,
                             R_avg_DISTANCE_landed, R_avg_opp_DISTANCE_att, 
                             R_avg_opp_DISTANCE_landed, R_avg_CLINCH_att, 
                             R_avg_CLINCH_landed, R_avg_opp_CLINCH_att, 
                             R_avg_opp_CLINCH_landed, R_avg_GROUND_att, R_avg_GROUND_landed,
                             R_avg_opp_GROUND_att, R_avg_opp_GROUND_landed, R_avg_CTRL_time.seconds.,
                             R_avg_opp_CTRL_time.seconds., R_total_time_fought.seconds., 
                             R_total_rounds_fought, R_total_title_bouts, R_current_win_streak, 
                             R_current_lose_streak, R_longest_win_streak, R_wins, R_losses, R_draw,
                             R_win_by_Decision_Majority, R_win_by_Decision_Split,
                             R_win_by_Decision_Unanimous, R_win_by_KO.TKO, R_win_by_Submission,
                             R_win_by_TKO_Doctor_Stoppage, R_Stance, R_Height_cms, R_Reach_cms, R_Weight_lbs,
                             R_age, B_fighter, B_avg_KD, B_avg_opp_KD, B_avg_SIG_STR_pct, 
                             B_avg_opp_SIG_STR_pct, B_avg_TD_pct, B_avg_opp_TD_pct, B_avg_SUB_ATT, 
                             B_avg_opp_SUB_ATT, B_avg_REV, B_avg_opp_REV, B_avg_SIG_STR_att, 
                             B_avg_SIG_STR_landed, B_avg_opp_SIG_STR_att,
                             B_avg_opp_SIG_STR_landed, B_avg_TOTAL_STR_att, B_avg_TOTAL_STR_landed,
                             B_avg_opp_TOTAL_STR_att, B_avg_opp_TOTAL_STR_landed, B_avg_TD_att,
                             B_avg_TD_landed, B_avg_opp_TD_att, B_avg_opp_TD_landed, B_avg_HEAD_att,
                             B_avg_HEAD_landed, B_avg_opp_HEAD_att, B_avg_opp_HEAD_landed, B_avg_BODY_att,
                             B_avg_BODY_landed, B_avg_opp_BODY_att, B_avg_opp_BODY_landed, B_avg_LEG_att,
                             B_avg_LEG_landed, B_avg_opp_LEG_att, B_avg_opp_LEG_landed, B_avg_DISTANCE_att,
                             B_avg_DISTANCE_landed, B_avg_opp_DISTANCE_att, B_avg_opp_DISTANCE_landed,
                             B_avg_CLINCH_att, B_avg_CLINCH_landed, B_avg_opp_CLINCH_att,
                             B_avg_opp_CLINCH_landed, B_avg_GROUND_att, B_avg_GROUND_landed,
                             B_avg_opp_GROUND_att, B_avg_opp_GROUND_landed, B_avg_CTRL_time.seconds.,
                             B_avg_opp_CTRL_time.seconds., B_total_time_fought.seconds.,
                             B_total_rounds_fought, B_total_title_bouts, B_current_win_streak,
                             B_current_lose_streak, B_longest_win_streak, B_wins, B_losses, B_draw,
                             B_win_by_Decision_Majority, B_win_by_Decision_Split,
                             B_win_by_Decision_Unanimous, B_win_by_KO.TKO, B_win_by_Submission,
                             B_win_by_TKO_Doctor_Stoppage, B_Stance, B_Height_cms, B_Reach_cms,
                             B_Weight_lbs, B_age))
  }
}
```

```{r}
ufc_data_shuffled <- data2 %>%
  rowwise() %>%
  mutate(shuffled_data = list(shuffle_fighters(B_fighter, B_avg_KD, B_avg_opp_KD, B_avg_SIG_STR_pct,
                                               B_avg_opp_SIG_STR_pct,B_avg_TD_pct,B_avg_opp_TD_pct, B_avg_SUB_ATT,
                                               B_avg_opp_SUB_ATT, B_avg_REV, B_avg_opp_REV, 
                                               B_avg_SIG_STR_att, B_avg_SIG_STR_landed,
                                               B_avg_opp_SIG_STR_att,B_avg_opp_SIG_STR_landed, 
                                               B_avg_TOTAL_STR_att, B_avg_TOTAL_STR_landed,
                                               B_avg_opp_TOTAL_STR_att, B_avg_opp_TOTAL_STR_landed, 
                                               B_avg_TD_att, B_avg_TD_landed, B_avg_opp_TD_att, 
                                               B_avg_opp_TD_landed, B_avg_HEAD_att,
                                               B_avg_HEAD_landed, B_avg_opp_HEAD_att, 
                                               B_avg_opp_HEAD_landed, B_avg_BODY_att,
                                               B_avg_BODY_landed, B_avg_opp_BODY_att, 
                                               B_avg_opp_BODY_landed, B_avg_LEG_att, 
                                               B_avg_LEG_landed, B_avg_opp_LEG_att, 
                                               B_avg_opp_LEG_landed, B_avg_DISTANCE_att,
                                               B_avg_DISTANCE_landed, B_avg_opp_DISTANCE_att, 
                                               B_avg_opp_DISTANCE_landed,B_avg_CLINCH_att, 
                                               B_avg_CLINCH_landed, B_avg_opp_CLINCH_att,
                                               B_avg_opp_CLINCH_landed, B_avg_GROUND_att, B_avg_GROUND_landed,
                                               B_avg_opp_GROUND_att, B_avg_opp_GROUND_landed, B_avg_CTRL_time.seconds.,
                                               B_avg_opp_CTRL_time.seconds., B_total_time_fought.seconds.,
                                               B_total_rounds_fought, B_total_title_bouts, B_current_win_streak,
                                               B_current_lose_streak, B_longest_win_streak, B_wins, B_losses, B_draw,
                                               B_win_by_Decision_Majority, B_win_by_Decision_Split,
                                               B_win_by_Decision_Unanimous, B_win_by_KO.TKO, B_win_by_Submission,
                                               B_win_by_TKO_Doctor_Stoppage, B_Stance, B_Height_cms, B_Reach_cms,
                                               B_Weight_lbs, B_age, R_fighter, R_avg_KD, R_avg_opp_KD, 
                                               R_avg_SIG_STR_pct, 
                                               R_avg_opp_SIG_STR_pct, R_avg_TD_pct, R_avg_opp_TD_pct, R_avg_SUB_ATT, 
                                               R_avg_opp_SUB_ATT, R_avg_REV, R_avg_opp_REV, R_avg_SIG_STR_att, 
                                               R_avg_SIG_STR_landed, R_avg_opp_SIG_STR_att,
                                               R_avg_opp_SIG_STR_landed, R_avg_TOTAL_STR_att, R_avg_TOTAL_STR_landed,
                                               R_avg_opp_TOTAL_STR_att, R_avg_opp_TOTAL_STR_landed, R_avg_TD_att,
                                               R_avg_TD_landed, R_avg_opp_TD_att, R_avg_opp_TD_landed, R_avg_HEAD_att,
                                               R_avg_HEAD_landed, R_avg_opp_HEAD_att, R_avg_opp_HEAD_landed,
                                               R_avg_BODY_att,
                                               R_avg_BODY_landed, R_avg_opp_BODY_att, R_avg_opp_BODY_landed,
                                               R_avg_LEG_att,
                                               R_avg_LEG_landed, R_avg_opp_LEG_att, R_avg_opp_LEG_landed,
                                               R_avg_DISTANCE_att,
                                               R_avg_DISTANCE_landed, R_avg_opp_DISTANCE_att, R_avg_opp_DISTANCE_landed,
                                               R_avg_CLINCH_att, R_avg_CLINCH_landed, R_avg_opp_CLINCH_att,
                                               R_avg_opp_CLINCH_landed, R_avg_GROUND_att, R_avg_GROUND_landed,
                                               R_avg_opp_GROUND_att, R_avg_opp_GROUND_landed, R_avg_CTRL_time.seconds.,
                                               R_avg_opp_CTRL_time.seconds., R_total_time_fought.seconds.,
                                               R_total_rounds_fought, R_total_title_bouts, R_current_win_streak,
                                               R_current_lose_streak, R_longest_win_streak, R_wins, R_losses, R_draw,
                                               R_win_by_Decision_Majority, R_win_by_Decision_Split,
                                               R_win_by_Decision_Unanimous, R_win_by_KO.TKO, R_win_by_Submission,
                                               R_win_by_TKO_Doctor_Stoppage, R_Stance, R_Height_cms, R_Reach_cms,
                                               R_Weight_lbs, R_age))) %>%
  ungroup()
```


```{r}

ufc_data_shuffled <- ufc_data_shuffled %>%
  mutate(B_fighter = map_chr(shuffled_data, 1),
         B_avg_KD = map_chr(shuffled_data, 2),
         B_avg_opp_KD = map_chr(shuffled_data, 3),
         B_avg_SIG_STR_pct = map_chr(shuffled_data, 4),
         B_avg_opp_SIG_STR_pct = map_chr(shuffled_data, 5),
         B_avg_TD_pct = map_chr(shuffled_data, 6),
         B_avg_opp_TD_pct = map_chr(shuffled_data, 7),
         B_avg_SUB_ATT = map_chr(shuffled_data, 8),
         B_avg_opp_SUB_ATT = map_chr(shuffled_data, 9),
         B_avg_REV = map_chr(shuffled_data, 10),
         B_avg_opp_REV = map_chr(shuffled_data, 11),
         B_avg_SIG_STR_att = map_chr(shuffled_data, 12),
         B_avg_SIG_STR_landed = map_chr(shuffled_data, 13),
         B_avg_opp_SIG_STR_att = map_chr(shuffled_data, 14),
         B_avg_opp_SIG_STR_landed = map_chr(shuffled_data, 15),
         B_avg_TOTAL_STR_att = map_chr(shuffled_data, 16),
         B_avg_TOTAL_STR_landed = map_chr(shuffled_data, 17),
         B_avg_opp_TOTAL_STR_att = map_chr(shuffled_data, 18),
         B_avg_opp_TOTAL_STR_landed = map_chr(shuffled_data, 19),
         B_avg_TD_att = map_chr(shuffled_data, 20),
         B_avg_TD_landed = map_chr(shuffled_data, 21),
         B_avg_opp_TD_att = map_chr(shuffled_data, 22),
         B_avg_opp_TD_landed = map_chr(shuffled_data, 23),
         B_avg_HEAD_att = map_chr(shuffled_data, 24),
         B_avg_HEAD_landed = map_chr(shuffled_data, 25),
         B_avg_opp_HEAD_att = map_chr(shuffled_data, 26),
         B_avg_opp_HEAD_landed = map_chr(shuffled_data, 27),
         B_avg_BODY_att = map_chr(shuffled_data, 28),
         B_avg_BODY_landed = map_chr(shuffled_data, 29),
         B_avg_opp_BODY_att = map_chr(shuffled_data, 30),
         B_avg_opp_BODY_landed = map_chr(shuffled_data, 31),
         B_avg_LEG_att = map_chr(shuffled_data, 32),
         B_avg_LEG_landed = map_chr(shuffled_data, 33),
         B_avg_opp_LEG_att = map_chr(shuffled_data, 34),
         B_avg_opp_LEG_landed = map_chr(shuffled_data, 35),
         B_avg_DISTANCE_att = map_chr(shuffled_data, 36),
         B_avg_DISTANCE_landed = map_chr(shuffled_data, 37),
         B_avg_opp_DISTANCE_att = map_chr(shuffled_data, 38),
         B_avg_opp_DISTANCE_landed = map_chr(shuffled_data, 39),
         B_avg_CLINCH_att = map_chr(shuffled_data, 40),
         B_avg_CLINCH_landed = map_chr(shuffled_data, 41),
         B_avg_opp_CLINCH_att = map_chr(shuffled_data, 42),
         B_avg_opp_CLINCH_landed = map_chr(shuffled_data, 43),
         B_avg_GROUND_att = map_chr(shuffled_data, 44),
         B_avg_GROUND_landed = map_chr(shuffled_data, 45),
         B_avg_opp_GROUND_att = map_chr(shuffled_data, 46),
         B_avg_opp_GROUND_landed = map_chr(shuffled_data, 47),
         B_avg_CTRL_time.seconds. = map_chr(shuffled_data, 48),
         B_avg_opp_CTRL_time.seconds. = map_chr(shuffled_data, 49),
         B_total_time_fought.seconds. = map_chr(shuffled_data, 50),
         B_total_rounds_fought = map_chr(shuffled_data, 51),
         B_total_title_bouts = map_chr(shuffled_data, 52),
         B_current_win_streak = map_chr(shuffled_data, 53),
         B_current_lose_streak = map_chr(shuffled_data, 54),
         B_longest_win_streak = map_chr(shuffled_data, 55),
         B_wins = map_chr(shuffled_data, 56),
         B_losses = map_chr(shuffled_data, 57),
         B_draw = map_chr(shuffled_data, 58),
         B_win_by_Decision_Majority = map_chr(shuffled_data, 59),
         B_win_by_Decision_Split = map_chr(shuffled_data, 60),
         B_win_by_Decision_Unanimous = map_chr(shuffled_data, 61),
         B_win_by_KO.TKO = map_chr(shuffled_data, 62),
         B_win_by_Submission = map_chr(shuffled_data, 63),
         B_win_by_TKO_Doctor_Stoppage = map_chr(shuffled_data, 64),
         B_Stance = map_chr(shuffled_data, 65),
         B_Height_cms = map_chr(shuffled_data, 66),
         B_Reach_cms = map_chr(shuffled_data, 67),
         B_Weight_lbs = map_chr(shuffled_data, 68),
         B_age = map_chr(shuffled_data, 69),
         R_fighter = map_chr(shuffled_data, 70),
         R_avg_KD = map_chr(shuffled_data, 71),
         R_avg_opp_KD = map_chr(shuffled_data, 72),
         R_avg_SIG_STR_pct = map_chr(shuffled_data, 73),
         R_avg_opp_SIG_STR_pct = map_chr(shuffled_data, 74),
         R_avg_TD_pct = map_chr(shuffled_data, 75),
         R_avg_opp_TD_pct = map_chr(shuffled_data, 76),
         R_avg_SUB_ATT = map_chr(shuffled_data, 77),
         R_avg_opp_SUB_ATT = map_chr(shuffled_data, 78),
         R_avg_REV = map_chr(shuffled_data, 79),
         R_avg_opp_REV = map_chr(shuffled_data, 80),
         R_avg_SIG_STR_att = map_chr(shuffled_data, 81),
         R_avg_SIG_STR_landed = map_chr(shuffled_data, 82),
         R_avg_opp_SIG_STR_att = map_chr(shuffled_data, 83),
         R_avg_opp_SIG_STR_landed = map_chr(shuffled_data, 84),
         R_avg_TOTAL_STR_att = map_chr(shuffled_data, 85),
         R_avg_TOTAL_STR_landed = map_chr(shuffled_data, 86),
         R_avg_opp_TOTAL_STR_att = map_chr(shuffled_data, 87),
         R_avg_opp_TOTAL_STR_landed = map_chr(shuffled_data, 88),
         R_avg_TD_att = map_chr(shuffled_data, 89),
         R_avg_TD_landed = map_chr(shuffled_data, 90),
         R_avg_opp_TD_att = map_chr(shuffled_data, 91),
         R_avg_opp_TD_landed = map_chr(shuffled_data, 92),
         R_avg_HEAD_att = map_chr(shuffled_data, 93),
         R_avg_HEAD_landed = map_chr(shuffled_data, 94),
         R_avg_opp_HEAD_att = map_chr(shuffled_data, 95),
         R_avg_opp_HEAD_landed = map_chr(shuffled_data, 96),
         R_avg_BODY_att = map_chr(shuffled_data, 97),
         R_avg_BODY_landed = map_chr(shuffled_data, 98),
         R_avg_opp_BODY_att = map_chr(shuffled_data, 99),
         R_avg_opp_BODY_landed = map_chr(shuffled_data, 100),
         R_avg_LEG_att = map_chr(shuffled_data, 101),
         R_avg_LEG_landed = map_chr(shuffled_data, 102),
         R_avg_opp_LEG_att = map_chr(shuffled_data, 103),
         R_avg_opp_LEG_landed = map_chr(shuffled_data, 104),
         R_avg_DISTANCE_att = map_chr(shuffled_data, 105),
         R_avg_DISTANCE_landed = map_chr(shuffled_data, 106),
         R_avg_opp_DISTANCE_att = map_chr(shuffled_data, 107),
         R_avg_opp_DISTANCE_landed = map_chr(shuffled_data, 108),
         R_avg_CLINCH_att = map_chr(shuffled_data, 109),
         R_avg_CLINCH_landed = map_chr(shuffled_data, 110),
         R_avg_opp_CLINCH_att = map_chr(shuffled_data, 111),
         R_avg_opp_CLINCH_landed = map_chr(shuffled_data, 112),
         R_avg_GROUND_att = map_chr(shuffled_data, 113),
         R_avg_GROUND_landed = map_chr(shuffled_data, 114),
         R_avg_opp_GROUND_att = map_chr(shuffled_data, 115),
         R_avg_opp_GROUND_landed = map_chr(shuffled_data, 116),
         R_avg_CTRL_time.seconds. = map_chr(shuffled_data, 117),
         R_avg_opp_CTRL_time.seconds. = map_chr(shuffled_data, 118),
         R_total_time_fought.seconds. = map_chr(shuffled_data, 119),
         R_total_rounds_fought = map_chr(shuffled_data, 120),
         R_total_title_bouts = map_chr(shuffled_data, 121),
         R_current_win_streak = map_chr(shuffled_data, 122),
         R_current_lose_streak = map_chr(shuffled_data, 123),
         R_longest_win_streak = map_chr(shuffled_data, 124),
         R_wins = map_chr(shuffled_data, 125),
         R_losses = map_chr(shuffled_data, 126),
         R_draw = map_chr(shuffled_data, 127),
         R_win_by_Decision_Majority = map_chr(shuffled_data, 128),
         R_win_by_Decision_Split = map_chr(shuffled_data, 129),
         R_win_by_Decision_Unanimous = map_chr(shuffled_data, 130),
         R_win_by_KO.TKO = map_chr(shuffled_data, 131),
         R_win_by_Submission = map_chr(shuffled_data, 132),
         R_win_by_TKO_Doctor_Stoppage = map_chr(shuffled_data, 133),
         R_Stance = map_chr(shuffled_data, 134),
         R_Height_cms = map_chr(shuffled_data, 135),
         R_Reach_cms = map_chr(shuffled_data, 136),
         R_Weight_lbs = map_chr(shuffled_data, 137),
         R_age = map_chr(shuffled_data, 138))
```






```{r}
ufc_data_shuffled <- ufc_data_shuffled %>%
  mutate(Winner = ifelse(Winner2 == R_fighter, "Red", "Blue"))
```

```{r}
#turning all appropriate columns numeric
ufc_data_shuffled <- ufc_data_shuffled %>%
  mutate_at(9:71, as.numeric) %>% 
  mutate_at(73:138, as.numeric) %>% 
  mutate_at(140:144, as.numeric)
```




```{r}
#distribution of winner for each side
ggplot(data = ufc_data_shuffled, aes(x = Winner, fill = Winner)) + geom_bar() + scale_fill_manual(values = c("blue", "red"))
```




```{r}
ufc_data_shuffled <- ufc_data_shuffled %>% mutate(weight = case_when(weight_class == "Flyweight" ~ 125,
                                   weight_class == "Bantamweight" ~ 135,
                                   weight_class == "Featherweight" ~ 145,
                                   weight_class == "Lightweight" ~ 155,
                                   weight_class == "Welterweight" ~ 170,
                                   weight_class == "Middleweight" ~ 185,
                                   weight_class == "LightHeavyweight" ~ 205, 
                                   weight_class == "Heavyweight" ~ 265, 
                                   weight_class == "OpenWeight" ~ 300))

ufc_data_shuffled <- ufc_data_shuffled %>% mutate(new_class = case_when(weight_class == "Flyweight" ~ "Class 1",
                                   weight_class == "Bantamweight" ~ "Class 1",
                                   weight_class == "Featherweight" ~ "Class 1",
                                   weight_class == "Lightweight" ~ "Class 2",
                                   weight_class == "Welterweight" ~ "Class 2",
                                   weight_class == "Middleweight" ~ "Class 3",
                                   weight_class == "LightHeavyweight" ~ "Class 3", 
                                   weight_class == "Heavyweight" ~ "Class 4", 
                                   weight_class == "OpenWeight" ~ "Class 4"))

freq <- ufc_data_shuffled %>% group_by(new_class, weight_class) %>% 
  summarize(weight = median(weight),
            count = n()) %>% arrange(weight)

kable(freq, caption = "Frequency of Weight Class Fights")


```


```{r}
new_class_freq <- ufc_data_shuffled %>% group_by(new_class) %>% summarize(freq = n())

ggplot(data=new_class_freq, aes(x=new_class, y=freq)) +
  geom_bar(stat="identity", fill="steelblue")+
  theme_minimal() + labs(x = "New Class for Model Building", y = "Count", 
                         title = "Frequency Distribution of New Weight Classes")
```

```{r}
## EDA that justifies splitting model by weight class 
Ratt_class <- ufc_data_shuffled %>% group_by(new_class) %>% summarise(Rmean_head = mean(R_avg_HEAD_att, na.rm = TRUE), 
                                                        Rmean_body = mean(R_avg_BODY_att, na.rm = TRUE), 
                                                        Rmean_leg = mean(R_avg_LEG_att, na.rm = TRUE), 
                                                        Rmean_clinch = mean(R_avg_CLINCH_att, na.rm = TRUE),
                                                        Rmean_ground = mean(R_avg_GROUND_att, na.rm = TRUE),
                                                        Rmean_ctrltime = mean(R_avg_CTRL_time.seconds., na.rm = TRUE))
kable(Ratt_class, caption = "Avg Attempt of Different Style Attacks", digits = 1)


```



## Inference EDA/make model

# Removing Columns that will provide no meaning to our models
```{r}
ufc_data_shuffled_subsetted <-  ufc_data_shuffled[,!names(ufc_data_shuffled) %in% c("R_fighter", "B_fighter", "Referee",
                                                                                    "location", "first_fight", "Winner2",
                                                                                    "shuffled_data", "date", "fight_gender",
                                                                                    "year", "weight", "weight_class")]
```

```{r}
ufc_data_shuffled_subsetted  %>%
  group_by(R_draw) %>%
  summarise(count = n())

ufc_data_shuffled_subsetted  %>%
  group_by(B_draw) %>%
  summarise(count = n())

ufc_data_shuffled_subsetted <-  ufc_data_shuffled_subsetted[,!names(ufc_data_shuffled_subsetted) 
                                                            %in% c("R_draw", "B_draw")]
```

No draws. Going to remove this as well.


# checking for NA values
```{r}
any(is.na(ufc_data_shuffled_subsetted))
sum(is.na(ufc_data_shuffled_subsetted))
colSums(is.na(ufc_data_shuffled_subsetted))
```


# removing all rows with NA values
```{r}
ufc_data_shuffled_subsetted <- na.omit(ufc_data_shuffled_subsetted)

nrow(ufc_data_shuffled)
nrow(ufc_data_shuffled_subsetted)
nrow(ufc_data_shuffled) - nrow(ufc_data_shuffled_subsetted)
```


# creating 4 different dataframes, dividing up by our weight class division
```{r}
class1 <- ufc_data_shuffled_subsetted %>% filter(new_class == "Class 1")
class2 <- ufc_data_shuffled_subsetted %>% filter(new_class == "Class 2")
class3 <- ufc_data_shuffled_subsetted %>% filter(new_class == "Class 3")
class4 <- ufc_data_shuffled_subsetted %>% filter(new_class == "Class 4")

class1 <-  class1[,!names(class1) %in% c("new_class")]
class2 <-  class2[,!names(class2) %in% c("new_class")]
class3 <-  class3[,!names(class3) %in% c("new_class")]
class4 <-  class4[,!names(class4) %in% c("new_class")]

nrow(class1)
nrow(class2)
nrow(class3)
nrow(class4)
```


#check multicollinearity
```{r}
corr_data_class1 <- class1[sapply(class1, is.numeric)]
corr_data_class1 <- corr_data_class1[, sapply(corr_data_class1, function(x) sd(x, na.rm = TRUE) > 0)]

extract_high_corr <- function(corr_data_class1, threshold = 0.7) {
  corr_matrix <- cor(corr_data_class1, use = "complete.obs")
  high_corr <- data.frame(Var1 = character(), 
                                  Var2 = character(), 
                                  Correlation = numeric())
  for (i in 1:(ncol(corr_matrix)-1)) {
    for (j in (i+1):ncol(corr_matrix)) {
      if (abs(corr_matrix[i, j]) > threshold) {
        high_corr <- rbind(high_corr, 
                           data.frame(Var1 = colnames(corr_matrix)[i],
                                      Var2 = colnames(corr_matrix)[j],
                                      Correlation = corr_matrix[i, j]))
      }
    }
  }
  return(high_corr)
}

extract_high_corr(corr_data_class1, 0.7)
```

```{r}
corr_data_class2 <- class2[sapply(class2, is.numeric)]
corr_data_class2 <- corr_data_class2[, sapply(corr_data_class2, function(x) sd(x, na.rm = TRUE) > 0)]

extract_high_corr <- function(corr_data_class2, threshold = 0.7) {
  corr_matrix <- cor(corr_data_class2, use = "complete.obs")
  high_corr <- data.frame(Var1 = character(), 
                                  Var2 = character(), 
                                  Correlation = numeric())
  for (i in 1:(ncol(corr_matrix)-1)) {
    for (j in (i+1):ncol(corr_matrix)) {
      if (abs(corr_matrix[i, j]) > threshold) {
        high_corr <- rbind(high_corr, 
                           data.frame(Var1 = colnames(corr_matrix)[i],
                                      Var2 = colnames(corr_matrix)[j],
                                      Correlation = corr_matrix[i, j]))
      }
    }
  }
  return(high_corr)
}

extract_high_corr(corr_data_class2, 0.7)
```

```{r}
corr_data_class3 <- class3[sapply(class3, is.numeric)]
corr_data_class3 <- corr_data_class3[, sapply(corr_data_class3, function(x) sd(x, na.rm = TRUE) > 0)]

extract_high_corr <- function(corr_data_class3, threshold = 0.7) {
  corr_matrix <- cor(corr_data_class3, use = "complete.obs")
  high_corr <- data.frame(Var1 = character(), 
                                  Var2 = character(), 
                                  Correlation = numeric())
  for (i in 1:(ncol(corr_matrix)-1)) {
    for (j in (i+1):ncol(corr_matrix)) {
      if (abs(corr_matrix[i, j]) > threshold) {
        high_corr <- rbind(high_corr, 
                           data.frame(Var1 = colnames(corr_matrix)[i],
                                      Var2 = colnames(corr_matrix)[j],
                                      Correlation = corr_matrix[i, j]))
      }
    }
  }
  return(high_corr)
}

extract_high_corr(corr_data_class3, 0.7)
```

```{r}
corr_data_class4 <- class4[sapply(class4, is.numeric)]
corr_data_class4 <- corr_data_class4[, sapply(corr_data_class4, function(x) sd(x, na.rm = TRUE) > 0)]

extract_high_corr <- function(corr_data_class4, threshold = 0.7) {
  corr_matrix <- cor(corr_data_class4, use = "complete.obs")
  high_corr <- data.frame(Var1 = character(), 
                                  Var2 = character(), 
                                  Correlation = numeric())
  for (i in 1:(ncol(corr_matrix)-1)) {
    for (j in (i+1):ncol(corr_matrix)) {
      if (abs(corr_matrix[i, j]) > threshold) {
        high_corr <- rbind(high_corr, 
                           data.frame(Var1 = colnames(corr_matrix)[i],
                                      Var2 = colnames(corr_matrix)[j],
                                      Correlation = corr_matrix[i, j]))
      }
    }
  }
  return(high_corr)
}

extract_high_corr(corr_data_class4, 0.7)
```

There are so many predictor variables (over 100), we need to reduce that number. First, creating correlation matrix to identify variables that are highly correlated with each other. It would be much to visualize all combinations, so we will only extract high correlation combos (correlation > 0.7), for each class dataframe.

over 180 combinations in every class have high correlation. A lot.

Now need to deal with multicollinearity and perform variable selection. Will use ridge regression (to reduce multicollinearity) and lasso (to select varialbes). This is know as Elastic Net (linear regression model that includes both the L1 penalty of Lasso and the L2 penalty of Ridge regression). Elastic Net combines both L2 and L1 penalties of ridge regression and lasso. It controls the mixing of the two penalties through a parameter (lambda).


#scale numerical predictors
```{r}
names(class1)[sapply(class1, is.numeric)]

numerical_columns <- c("B_avg_KD", "B_avg_opp_KD", "B_avg_SIG_STR_pct",
                       "B_avg_opp_SIG_STR_pct", "B_avg_TD_pct", "B_avg_opp_TD_pct", "B_avg_SUB_ATT",
                       "B_avg_opp_SUB_ATT", "B_avg_REV", "B_avg_opp_REV", 
                       "B_avg_SIG_STR_att", "B_avg_SIG_STR_landed",
                       "B_avg_opp_SIG_STR_att", "B_avg_opp_SIG_STR_landed", 
                       "B_avg_TOTAL_STR_att", "B_avg_TOTAL_STR_landed",
                       "B_avg_opp_TOTAL_STR_att", "B_avg_opp_TOTAL_STR_landed", 
                       "B_avg_TD_att", "B_avg_TD_landed", "B_avg_opp_TD_att", 
                       "B_avg_opp_TD_landed", "B_avg_HEAD_att",
                       "B_avg_HEAD_landed", "B_avg_opp_HEAD_att", 
                       "B_avg_opp_HEAD_landed", "B_avg_BODY_att",
                       "B_avg_BODY_landed", "B_avg_opp_BODY_att", 
                       "B_avg_opp_BODY_landed", "B_avg_LEG_att", 
                       "B_avg_LEG_landed", "B_avg_opp_LEG_att", 
                       "B_avg_opp_LEG_landed", "B_avg_DISTANCE_att",
                       "B_avg_DISTANCE_landed", "B_avg_opp_DISTANCE_att", 
                       "B_avg_opp_DISTANCE_landed", "B_avg_CLINCH_att", 
                       "B_avg_CLINCH_landed", "B_avg_opp_CLINCH_att",
                       "B_avg_opp_CLINCH_landed", "B_avg_GROUND_att", "B_avg_GROUND_landed",
                       "B_avg_opp_GROUND_att", "B_avg_opp_GROUND_landed", "B_avg_CTRL_time.seconds.",
                       "B_avg_opp_CTRL_time.seconds.", "B_total_time_fought.seconds.",
                       "B_total_rounds_fought", "B_total_title_bouts", "B_current_win_streak",
                       "B_current_lose_streak", "B_longest_win_streak", "B_wins", "B_losses",
                       "B_win_by_Decision_Majority", "B_win_by_Decision_Split",
                       "B_win_by_Decision_Unanimous", "B_win_by_KO.TKO", "B_win_by_Submission",
                       "B_win_by_TKO_Doctor_Stoppage", "B_Height_cms", "B_Reach_cms",
                       "B_Weight_lbs", "B_age", "R_avg_KD", "R_avg_opp_KD", "R_avg_SIG_STR_pct", 
                       "R_avg_opp_SIG_STR_pct", "R_avg_TD_pct", "R_avg_opp_TD_pct", "R_avg_SUB_ATT", 
                       "R_avg_opp_SUB_ATT", "R_avg_REV", "R_avg_opp_REV", "R_avg_SIG_STR_att", 
                       "R_avg_SIG_STR_landed", "R_avg_opp_SIG_STR_att",
                       "R_avg_opp_SIG_STR_landed", "R_avg_TOTAL_STR_att", "R_avg_TOTAL_STR_landed",
                       "R_avg_opp_TOTAL_STR_att", "R_avg_opp_TOTAL_STR_landed", "R_avg_TD_att",
                       "R_avg_TD_landed", "R_avg_opp_TD_att", "R_avg_opp_TD_landed", "R_avg_HEAD_att",
                       "R_avg_HEAD_landed", "R_avg_opp_HEAD_att", "R_avg_opp_HEAD_landed", "R_avg_BODY_att",
                       "R_avg_BODY_landed", "R_avg_opp_BODY_att", "R_avg_opp_BODY_landed", "R_avg_LEG_att",
                       "R_avg_LEG_landed", "R_avg_opp_LEG_att", "R_avg_opp_LEG_landed", "R_avg_DISTANCE_att",
                       "R_avg_DISTANCE_landed", "R_avg_opp_DISTANCE_att", "R_avg_opp_DISTANCE_landed",
                       "R_avg_CLINCH_att", "R_avg_CLINCH_landed", "R_avg_opp_CLINCH_att",
                       "R_avg_opp_CLINCH_landed", "R_avg_GROUND_att", "R_avg_GROUND_landed",
                       "R_avg_opp_GROUND_att", "R_avg_opp_GROUND_landed", "R_avg_CTRL_time.seconds.",
                       "R_avg_opp_CTRL_time.seconds.", "R_total_time_fought.seconds.",
                       "R_total_rounds_fought", "R_total_title_bouts", "R_current_win_streak",
                       "R_current_lose_streak", "R_longest_win_streak", "R_wins", "R_losses",
                       "R_win_by_Decision_Majority", "R_win_by_Decision_Split",
                       "R_win_by_Decision_Unanimous", "R_win_by_KO.TKO", "R_win_by_Submission",
                       "R_win_by_TKO_Doctor_Stoppage", "R_Height_cms", "R_Reach_cms", "R_Weight_lbs", "R_age")

class1_scaled <- class1
class1_scaled[, numerical_columns] <- scale(class1[, numerical_columns])

class2_scaled <- class2
class2_scaled[, numerical_columns] <- scale(class2[, numerical_columns])

class3_scaled <- class3
class3_scaled[, numerical_columns] <- scale(class3[, numerical_columns])

class4_scaled <- class4
class4_scaled[, numerical_columns] <- scale(class4[, numerical_columns])
```

I scaled the numerical predictors for each dataset because these techniques are sensitive to scale.


#transform categorical predictors
```{r}
library(caret)

dummies1 <- dummyVars(~ Winner + title_bout + B_Stance + R_Stance, data = class1_scaled)
dummy_data1 <- predict(dummies1, newdata = class1_scaled)
class1_scaled <- cbind(class1_scaled[, !names(class1_scaled) %in% c('Winner', 
                                                                    'title_bout', 'B_Stance', 'R_Stance')], dummy_data1)
class1_scaled <-  class1_scaled[,!names(class1_scaled) %in% c("WinnerBlue")]

dummies2 <- dummyVars(~ Winner + title_bout + B_Stance + R_Stance, data = class2_scaled)
dummy_data2 <- predict(dummies2, newdata = class2_scaled)
class2_scaled <- cbind(class2_scaled[, !names(class2_scaled) %in% c('Winner',
                                                                    'title_bout', 'B_Stance', 'R_Stance')], dummy_data2)
class2_scaled <-  class2_scaled[,!names(class2_scaled) %in% c("WinnerBlue")]

dummies3 <- dummyVars(~ Winner + title_bout + B_Stance + R_Stance, data = class3_scaled)
dummy_data3 <- predict(dummies3, newdata = class3_scaled)
class3_scaled <- cbind(class3_scaled[, !names(class3_scaled) %in% c('Winner',
                                                                    'title_bout', 'B_Stance', 'R_Stance')], dummy_data3)
class3_scaled <-  class3_scaled[,!names(class3_scaled) %in% c("WinnerBlue")]

dummies4 <- dummyVars(~ Winner + title_bout + B_Stance + R_Stance, data = class4_scaled)
dummy_data4 <- predict(dummies4, newdata = class4_scaled)
class4_scaled <- cbind(class4_scaled[, !names(class4_scaled) %in% c('Winner',
                                                                    'title_bout', 'B_Stance', 'R_Stance')], dummy_data4)
class4_scaled <-  class4_scaled[,!names(class4_scaled) %in% c("WinnerBlue")]
```

I will transform the categorical variables (including our response variable) into numerical format so they can be used in Elastic Net. I will do this by using dummy variables. I am removing the additional dummy variable for the response, so it is only in the context of Red winning or losing (1 = Red win, 0 = Red lose).


#create test and train datasets
```{r}
set.seed(2)
class1_scaled$id <- 1:nrow(class1_scaled)
class1_train <- class1_scaled %>% dplyr::sample_frac(0.7)
class1_test  <- dplyr::anti_join(class1_scaled, class1_train, by = 'id')

class2_scaled$id <- 1:nrow(class2_scaled)
class2_train <- class2_scaled %>% dplyr::sample_frac(0.7)
class2_test  <- dplyr::anti_join(class2_scaled, class2_train, by = 'id')

class3_scaled$id <- 1:nrow(class3_scaled)
class3_train <- class3_scaled %>% dplyr::sample_frac(0.7)
class3_test  <- dplyr::anti_join(class3_scaled, class3_train, by = 'id')

class4_scaled$id <- 1:nrow(class4_scaled)
class4_train <- class4_scaled %>% dplyr::sample_frac(0.7)
class4_test  <- dplyr::anti_join(class4_scaled, class4_train, by = 'id')


train_x1 <- as.matrix(class1_train[, !names(class1_train) %in% 'WinnerRed'])
train_y1 <- class1_train$WinnerRed
test_x1 <- as.matrix(class1_test[, !names(class1_test) %in% 'WinnerRed'])
test_y1 <- class1_test$WinnerRed

train_x2 <- as.matrix(class2_train[, !names(class2_train) %in% 'WinnerRed'])
train_y2 <- class2_train$WinnerRed
test_x2 <- as.matrix(class2_test[, !names(class2_test) %in% 'WinnerRed'])
test_y2 <- class2_test$WinnerRed

train_x3 <- as.matrix(class3_train[, !names(class3_train) %in% 'WinnerRed'])
train_y3 <- class3_train$WinnerRed
test_x3 <- as.matrix(class3_test[, !names(class3_test) %in% 'WinnerRed'])
test_y3 <- class3_test$WinnerRed

train_x4 <- as.matrix(class4_train[, !names(class4_train) %in% 'WinnerRed'])
train_y4 <- class4_train$WinnerRed
test_x4 <- as.matrix(class4_test[, !names(class4_test) %in% 'WinnerRed'])
test_y4 <- class4_test$WinnerRed
```

use 70% of dataset as training set and 30% as test set. separate the response variable from the predictor variables


#final model (class 1)
```{r}
library(glmnet)
set.seed(2)

alpha_values <- seq(0, 1, by = 0.1)
cv_results1 <- lapply(alpha_values, function(alpha) {
  cv.glmnet(train_x1, train_y1, alpha = alpha, family = "binomial", type.measure = "deviance")
})

best_cv1 <- sapply(cv_results1, function(cv) min(cv$cvm))
best_alpha_index1 <- which.min(best_cv1)
best_alpha1 <- alpha_values[best_alpha_index1]
best_lambda1 <- cv_results1[[best_alpha_index1]]$lambda.min

final_model1 <- glmnet(train_x1, train_y1, alpha = best_alpha1, lambda = best_lambda1, family = "binomial")

predictions1 <- predict(final_model1, newx = test_x1, type = "response", s = best_lambda1)
predicted_classes1 <- ifelse(predictions1 > 0.5, 1, 0)
```

```{r}
library(e1071)

conf_matrix1 <- confusionMatrix(as.factor(predicted_classes1), as.factor(test_y1))
accuracy1 <- conf_matrix1$overall['Accuracy']
precision1 <- conf_matrix1$byClass['Precision']
recall1 <- conf_matrix1$byClass['Recall']
F1_Score1 <- conf_matrix1$byClass['F1']

accuracy1
precision1
recall1
F1_Score1
coef(final_model1, s = best_lambda1)
```

Now moving to models, perform cross-validation to find best lambda value. Will use to find the optimal lambda and alpha parameters. Alpha is the mixing parameter between Lasso (alpha = 1) and Ridge (alpha = 0) regression. Run cross-validation for each alpha value to find the optimal lambda. Find the combination that gives the best performance (e.g., the lowest deviance). Fit the Elastic Net model using the optimal alpha and lambda values. Evaluate the model on test set to check its performance.

Accuracy: proportion of the total number of predictions that were correct.
Precision: ratio of correctly predicted positive observations to the total predicted positives.
Recall (Sensitivity): ratio of correctly predicted positive observations to all observations in the actual class. 
F1-Score: weighted average of Precision and Recall.

Repeat with classes 2-4


#final model (class 2)
```{r}
set.seed(2)

alpha_values <- seq(0, 1, by = 0.1)
cv_results2 <- lapply(alpha_values, function(alpha) {
  cv.glmnet(train_x2, train_y2, alpha = alpha, family = "binomial", type.measure = "deviance")
})

best_cv2 <- sapply(cv_results2, function(cv) min(cv$cvm))
best_alpha_index2 <- which.min(best_cv2)
best_alpha2 <- alpha_values[best_alpha_index2]
best_lambda2 <- cv_results2[[best_alpha_index2]]$lambda.min

final_model2 <- glmnet(train_x2, train_y2, alpha = best_alpha2, lambda = best_lambda2, family = "binomial")

predictions2 <- predict(final_model2, newx = test_x2, type = "response", s = best_lambda2)
predicted_classes2 <- ifelse(predictions2 > 0.5, 1, 0)
```

```{r}
conf_matrix2 <- confusionMatrix(as.factor(predicted_classes2), as.factor(test_y2))
accuracy2 <- conf_matrix2$overall['Accuracy']
precision2 <- conf_matrix2$byClass['Precision']
recall2 <- conf_matrix2$byClass['Recall']
F1_Score2 <- conf_matrix2$byClass['F1']

accuracy2
precision2
recall2
F1_Score2
coef(final_model2, s = best_lambda2)
```


#final model (class 3)
```{r}
set.seed(2)

alpha_values <- seq(0, 1, by = 0.1)
cv_results3 <- lapply(alpha_values, function(alpha) {
  cv.glmnet(train_x3, train_y3, alpha = alpha, family = "binomial", type.measure = "deviance")
})

best_cv3 <- sapply(cv_results3, function(cv) min(cv$cvm))
best_alpha_index3 <- which.min(best_cv3)
best_alpha3 <- alpha_values[best_alpha_index3]
best_lambda3 <- cv_results3[[best_alpha_index3]]$lambda.min

final_model3 <- glmnet(train_x3, train_y3, alpha = best_alpha3, lambda = best_lambda3, family = "binomial")

predictions3 <- predict(final_model3, newx = test_x3, type = "response", s = best_lambda3)
predicted_classes3 <- ifelse(predictions3 > 0.5, 1, 0)
```

```{r}
conf_matrix3 <- confusionMatrix(as.factor(predicted_classes3), as.factor(test_y3))
accuracy3 <- conf_matrix3$overall['Accuracy']
precision3 <- conf_matrix3$byClass['Precision']
recall3 <- conf_matrix3$byClass['Recall']
F1_Score3 <- conf_matrix3$byClass['F1']

accuracy3
precision3
recall3
F1_Score3
coef(final_model3, s = best_lambda3)
```


#final model (class 4)
```{r}
set.seed(2)

alpha_values <- seq(0, 1, by = 0.1)
cv_results4 <- lapply(alpha_values, function(alpha) {
  cv.glmnet(train_x4, train_y4, alpha = alpha, family = "binomial", type.measure = "deviance")
})

best_cv4 <- sapply(cv_results4, function(cv) min(cv$cvm))
best_alpha_index4 <- which.min(best_cv4)
best_alpha4 <- alpha_values[best_alpha_index4]
best_lambda4 <- cv_results4[[best_alpha_index4]]$lambda.min

final_model4 <- glmnet(train_x4, train_y4, alpha = best_alpha4, lambda = best_lambda4, family = "binomial")

predictions4 <- predict(final_model4, newx = test_x4, type = "response", s = best_lambda4)
predicted_classes4 <- ifelse(predictions4 > 0.5, 1, 0)
```

```{r}
conf_matrix4 <- confusionMatrix(as.factor(predicted_classes4), as.factor(test_y4))
accuracy4 <- conf_matrix4$overall['Accuracy']
precision4 <- conf_matrix4$byClass['Precision']
recall4 <- conf_matrix4$byClass['Recall']
F1_Score4 <- conf_matrix4$byClass['F1']

accuracy4
precision4
recall4
F1_Score4
coef(final_model4, s = best_lambda4)
```







